\documentclass{article}

\usepackage[a4paper,margin=24mm]{geometry}
\usepackage{array}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{hyperref}
\hypersetup{
  pdftitle={The pongtex Package},
  pdfauthor={Lilia Chen May Wither},
  pdfsubject={Typesetting Riichi Mahjong hands and discard rivers},
  pdfkeywords={LaTeX, Mahjong, Riichi, MPSZ, CTAN}
}
\usepackage{pongtex}

\setlength\parindent{0pt}
\setlength\parskip{6pt}
\newcommand\pkg[1]{\textsf{#1}}
\newcommand\cmd[1]{\texttt{\string#1}}
\newcommand\meta[1]{\ensuremath{\langle}\textit{#1}\ensuremath{\rangle}}
\newcommand\code[1]{\texttt{\detokenize{#1}}}

\title{The \pkg{pongtex} Package\\\large Typesetting Riichi Mahjong Hands and Discard Rivers}
\author{Lilia Chen May Wither}
\date{Version 1.7.0, 2026-07-01}

\begin{document}
\maketitle

\begin{abstract}
\pkg{pongtex} typesets Japanese/Riichi Mahjong tiles from compact MPSZ notation.
It provides a document command, \cmd{\mahjong}, for hands and melds, and
\cmd{\mahjongriver} for discard rivers.  It supports red fives, face-down and
blank tiles, rotated and stacked tiles, concealed-kong shorthand, local and
global size options, and recolouring of tile backs.
\end{abstract}

\section{Loading the package}

A minimal document looks like this.



If you are working from the CTAN source archive, generate the runtime package file first:

\begin{verbatim}
latex pongtex.ins
\end{verbatim}

The distributed user manual lives in \texttt{doc/pongtex-doc.pdf}; its source is
\texttt{doc/pongtex-doc.tex}.  The package implementation source is
\texttt{pongtex.dtx}, and \texttt{pongtex.ins} extracts \texttt{pongtex.sty}.

\begin{verbatim}
\documentclass{article}
\usepackage{pongtex}
\begin{document}
\mahjong{111m456s111p11122z}
\end{document}
\end{verbatim}

The package expects the PDF artwork in the \texttt{tiles/} directory.  With the
default setting, \texttt{pongtex.sty} and \texttt{tiles/} should be installed in
the same TeX-searchable package directory.

\section{Package options}

Options can be supplied when loading the package or changed later with
\cmd{\pongtexsetup}.

\begin{verbatim}
\usepackage[
  height=1.5\baselineskip,
  scale=0.75,
  color=blue!70!black,
  no-aka=0
]{pongtex}

\pongtexsetup{height=1.5\baselineskip,scale=0.75,color=teal!65!black}
\end{verbatim}

\begin{longtable}{@{}p{0.25\linewidth}p{0.22\linewidth}p{0.45\linewidth}@{}}
\toprule
Option & Default & Description \\
\midrule
\texttt{height} & \verb|\baselineskip| & Height of one upright tile. \\
\texttt{scale} & \texttt{0.75} & Scale factor for the tile symbol relative to the tile face. \\
\texttt{tile-dir} & \texttt{tiles} & Directory containing the tile PDF files. \\
\texttt{color} & \texttt{none} & Recolour the tile back.  Use any \pkg{xcolor} colour expression, or \texttt{none} for the embedded artwork. \\
\texttt{no-aka} & \texttt{0} & Replace red fives \texttt{0m}, \texttt{0p}, and \texttt{0s} with regular fives when set to \texttt{1}. \\
\texttt{river-cols} & \texttt{6} & Number of tiles per discard-river row. \\
\texttt{river-row-gap} & \texttt{0pt} & Vertical gap between discard-river rows. \\
\bottomrule
\end{longtable}

\section{The \cmd{\mahjong} command}

\begin{verbatim}
\mahjong{<tiles>}
\mahjong[<height>][<scale>][<back-colour>][<local-keys>]{<tiles>}
\mahjong[<local-keys>]{<tiles>}
\end{verbatim}

The first optional argument also accepts a key-value list.  This allows compact
local overrides such as:

\begin{verbatim}
\mahjong[height=1.5\baselineskip,scale=0.75,color=teal!65!black,no-aka=1]{x 0m0p0s}
\end{verbatim}

\section{MPSZ notation}

A tile is written as one or more digits followed by a suit letter.  The suits are
\texttt{m} for manzu, \texttt{p} for pinzu, \texttt{s} for souzu, and \texttt{z}
for honors.

\begin{longtable}{@{}p{0.32\linewidth}p{0.58\linewidth}@{}}
\toprule
Input & Output \\
\midrule
\code{111m456s111p11122z} & \mahjong{111m456s111p11122z} \\
\code{1112345678999p} & \mahjong{1112345678999p} \\
\code{19m19s19p1234567z} & \mahjong{19m19s19p1234567z} \\
\bottomrule
\end{longtable}

\subsection{Special tokens}

\begin{longtable}{@{}p{0.18\linewidth}p{0.72\linewidth}@{}}
\toprule
Token & Meaning \\
\midrule
\texttt{0m}, \texttt{0p}, \texttt{0s} & Red fives: \mahjong{0m0p0s} \\
\texttt{x} & Face-down tile: \mahjong{x} \\
\texttt{?} & Unknown blank tile: \mahjong{?} \\
\texttt{-} & Full visual gap between groups. \\
\texttt{N-} & Proportional gap of \texttt{N/7} of one tile width, for example \texttt{2-}. \\
\bottomrule
\end{longtable}

\subsection{Rotated and stacked tiles}

The marker \texttt{*} or an apostrophe rotates the preceding tile sideways.  The
marker \texttt{+} or a double quote stacks two sideways copies of the preceding
tile, which is useful for kan notation.

\begin{longtable}{@{}p{0.36\linewidth}p{0.54\linewidth}@{}}
\toprule
Input & Output \\
\midrule
\code{111m111s111p22z2-3*333z} & \mahjong{111m111s111p22z2-3*333z} \\
\code{123m-xx?4p'5p"0s} & \mahjong{123m-xx?4p'5p"0s} \\
\bottomrule
\end{longtable}

\subsection{Red fives and the \texttt{no-aka} option}

Red fives are enabled by default.

\begin{longtable}{@{}p{0.36\linewidth}p{0.54\linewidth}@{}}
\toprule
Input & Output \\
\midrule
\code{0m0p0s55555m} & \mahjong{0m0p0s55555m} \\
\code{\mahjong[no-aka=1]{0m0p0s55555p}} & \mahjong[no-aka=1]{0m0p0s55555p} \\
\bottomrule
\end{longtable}

\subsection{Concealed kongs}

Five identical consecutive digits followed by a suit are rendered as a concealed
kan: a face-down tile, two visible tiles, and a face-down tile.  For suited fives,
the visible pair contains a red five and a regular five unless \texttt{no-aka=1}
is active.

\begin{longtable}{@{}p{0.36\linewidth}p{0.54\linewidth}@{}}
\toprule
Input & Output \\
\midrule
\code{55555p} & \mahjong{55555p} \\
\code{33333z} & \mahjong{33333z} \\
\code{\mahjong[no-aka=1]{55555s}} & \mahjong[no-aka=1]{55555s} \\
\bottomrule
\end{longtable}

\section{Tile-back recolouring}

The tile-back colour can be set globally or locally.  The colour value is passed
to \pkg{xcolor}.  The special value \texttt{none} restores the embedded back tile.

\begin{verbatim}
\pongtexsetup{color=purple!70!black}
\mahjong{x x x}

\mahjong[color=teal!65!black]{x x x}
\mahjong[color=none]{x x x}
\end{verbatim}

\begin{center}
\pongtexsetup{color=purple!70!black}
\mahjong{x x x}\quad
\mahjong[color=teal!65!black]{x x x}\quad
\mahjong[color=none]{x x x}
\end{center}
\pongtexsetup{color=none}

\section{Discard rivers}

The \cmd{\mahjongriver} command uses the same notation as \cmd{\mahjong}, but it
breaks tiles into rows.  The row length is controlled by \texttt{river-cols}.

\begin{verbatim}
\mahjongriver{1m9m2z5z3s6s1s1m8p4z87m}
\mahjongriver[river-cols=6]{1m9m2z5z3s6s1s*1m8p4z87m}
\end{verbatim}

\begin{center}
\mahjongriver{1m9m2z5z3s6s1s1m8p4z87m}

\mahjongriver[river-cols=6]{1m9m2z5z3s6s1s*1m8p4z87m}
\end{center}

\section{Expl3 interface}

For package authors who prefer an expl3-style interface, PongTeX also exposes:

\begin{verbatim}
\ExplSyntaxOn
\pongtex:n      {111m456s111p11122z}
\pongtexriver:n {1m2m3m4m5m6m7m8m9m}
\ExplSyntaxOff
\end{verbatim}

\section{Attribution and license}

The LaTeX package code is distributed under the MIT License.  Portions are based
on Daniel Schmitz's \pkg{mahjong} package, copyright 2021 Daniel Schmitz, also
under the MIT License.

The tile artwork in \texttt{tiles/} is derived from FluffyStuff's
\texttt{riichi-mahjong-tiles} project.  The upstream license file places that
work in the public domain/CC0.  Attribution is retained here and in the README.

\end{document}
